---
title: "Deliverable for Blue Water Baltimore"
author: "Eduardo Marin (with Dexter H Locke)"
date: "`r format(Sys.time())`"
output:
  html_document:
    theme: flatly
    code_folding: hide
    fig_width: 8
    fig_height: 7
    fig_caption: true
    toc: true
    toc_float: true
    self_contained: true
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

# Blue Water Baltimore & Tree Giveaways

An entry-level analysis from the past ten+ years (2013 to 2024)

The document was prepared by Dexter H. Locke and Eduardo Marin. The research document will include the following sections with accompanying downloadable visuals for all the graphs.
- Location & Time (2/2)
- Provider (1/1)
- Species (0/1)
- Watershed (0/1)
- Demographics (0/1)
- Origin-Destination (0/1)

```{r Loading Packages, message=FALSE, include=FALSE}
# Loading all packages

## List all of your packages here
knitr::opts_chunk$set(echo = TRUE)

packs <-c(
            'janitor'    # cleans things up, also pipe-friendly cross-tabulations
           , 'sf'         # for spatial data support
          , 'tidyverse'  # cuz
          # , 'tidylog'    # prints out what was done in dplyr and tidyr
          , 'magrittr'   # for the pipe
          , 'mapview'    # web maps for zooming and panning around
          #, 'beepr'      # makes noise when things are done!
          , 'tictoc'     # timing things.
          , 'raster'
          # , 'doParallel' # does what is says! PARALLEL
          # 'broom.mixed',# tidiers for mixed models AND nlme::gls()
          # , 'lubridate'   # DATES!
          , 'tidycensus' # tidy census package
          , 'tidygeocoder' # geo coding
          , 'leaflet' #creating the interactive mapping elements (more specific)
          , 'shiny'
          , 'leafsync'  # linked maps
          , 'RColorBrewer'
          , 'DT'
          , 'openxlsx'
          , 'mapdeck'
          , 'biscale' #used with ggplot to make bivariate maps
          , 'cowplot' #used to make more aesthetic ggplot visuals
          , 'ggspatial' #creates a basemap for your ggplots
          , 'classInt' #find me breaks without doing manualcalculations
          , 'rgl' #needs for rayshader
          , 'rayshader' #good data viz for 3d bar graphs
          , 'leaflet' #for heatmaps and more unique, customizable maps
          , 'leaflet.extras'
          )     

## IF the packages in 'packs' are not already installed, install them
## OTHERWISE do NOTHING
if (length(setdiff(packs, rownames(installed.packages()))) > 0) {
install.packages(setdiff(packs, rownames(installed.packages())))
}

## lapply(packs, library, character.only = TRUE)
## this actually loads them with library(package_name)
vapply(packs, library, character.only = TRUE, logical(1), logical.return = TRUE, quietly = TRUE)


## For tidycensus
## Setting for get_acs
census_api_key('58fc555c77c229747ade7d9fe50e7c71297cf91a', install = TRUE, overwrite = TRUE)
readRenviron("~/.Renviron")
options(tigris_use_cache = TRUE)
```

```{r Loading Census Data, include=FALSE}
# Pulling census data
## Listing needed variables
variables_acs <- 
  tidycensus::load_variables('acs5', year = 2021)

my_vars <- c(  'median_household_income' = 'B19013_001'
             , 'total_population' = 'B01003_001'
             , 'white_alone' = 'B03002_003'
             , 'median_age' = 'B01002_001'
             , 'total_population_over_25' = 'B15003_001'
             , 'ed_hs_degree' = 'B15003_017'
             , 'ed_ged_degree' = 'B15003_018'
             , 'ed_college_less_1_year' = 'B15003_019'
             , 'ed_college_more_1_year' = 'B15003_020'
             , 'ed_associate_degree' = 'B15003_021'
             , 'ed_plus_bachelor_degree' = 'B15003_022'
             , 'ed_plus_master_degree' = 'B15003_023'
             , 'ed_plus_professional_degree' = 'B15003_024'
             , 'ed_plus_doctorate_degree' = 'B15003_025'
            )

(maryland <-
  tidycensus::get_acs(
      geography = 'state'
    , state = 'MD'
    , variables = my_vars
    , year = 2021
    , geometry = TRUE
    , output = 'wide'
    , moe_level = 95
    ) |>
   st_transform(crs = st_crs(4326)))

(baltimore_city_counties <-
  tidycensus::get_acs(
      geography = 'county'
    , state = 'MD'
    , county = c('Baltimore City', 'Baltimore County')
    , variables = my_vars
    , year = 2021
    , geometry = TRUE
    , output = 'wide'
    , moe_level = 95
    ) |>
   st_transform(crs = st_crs(4326))) 

(city <-
  tidycensus::get_acs(
      geography = 'county'
    , state = 'MD'
    , county = 'Baltimore City'
    , variables = my_vars
    , year = 2021
    , geometry = TRUE
    , output = 'wide'
    , moe_level = 95
    ) |>
   st_transform(crs = st_crs(4326))) 

cbg <-
  tidycensus::get_acs(
      geography = 'block group'
    , state = 'Maryland'
    , county = c('Baltimore City', 'Baltimore County')
    , variables = my_vars
    , year = 2021
    , geometry = TRUE
    , output = 'wide'
    , moe_level = 95
    ) |>
    st_transform(crs = st_crs(4326)) |>
    mutate(jurisdiction = case_when(
      str_like(GEOID, '24510%') ~ "Baltimore City",
      str_like(GEOID, '24005%') ~ "Baltimore County",
      TRUE ~ "Neither"
    )) |>
    dplyr::select(-ends_with("M")) |> 
    rename_with(~ gsub("E$", "", .x), ends_with("E")) |>  
    rename(NAME = NAM) |>
    mutate(
      at_least_hs = rowSums(across(starts_with("ed_"))),
      at_least_college = rowSums(across(starts_with("ed_plus")))
    ) |>
    dplyr::select(-starts_with("ed_"))
```

```{r Loading Geographies, message=FALSE, include=FALSE}
# Reading in the shapefile and creating mapview layers
## Making the Maryland layer
state_layer <-
 maryland |>
 mapview(alpha.regions = 0, lwd = 2, col.regions = "#525252", alpha = 0.25, layer.name = "Maryland") 

## Making the Baltimore County layer
county_layer <-
  baltimore_city_counties |>
  mapview(alpha.regions = 0, lwd = 2, col.regions = "#525252", alpha = 0.25, layer.name = "Baltimore County") 

## Making the Baltimore city layer
city_layer <-
  city |>
  rename("Name" = "NAME") |>
  mutate("Name" = "Baltimore city") |>
  mapview(col.regions = "#525252", alpha = 0.25, layer.name = "Baltimore City") 

#Making the Baltimore neighborhood read
neighborhood_read <-
  st_read("../baltimore_tree_partners_data_to_big/shapefiles/Neighborhood/Neighborhood.shp") |>
  dplyr::select(`Name`,`geometry`,`Shape__Are`,`Shape__Len`) |>
  st_zm(drop = TRUE, what = "ZM") |>
  st_transform(4326) |>
  mutate(`Area (Square Miles)` = as.numeric(st_area(geometry)) / 2.59e+6)

## Making the Baltimore neighborhood layer
neighborhood_layer <-
  neighborhood_read |>
  mapview(alpha.regions = 0, lwd = 2, col.regions = "#525252", zcol = "Name", layer.name = "Baltimore's Neighborhood")

## Reading in the Blue Water Giveaway data
blue_water_giveaway_data <-
 st_read("../baltimore_tree_partners_data_to_big/shapefiles/bwb_giveaway_final/final_blue_water_baltimore_giveaways.shp")

## Making the Blue Water Giveaway layer
giveaway_layer <-
  blue_water_giveaway_data |>
  mapview(col.regions = "#5d6e75", alpha = 0.25, layer.name = "Total Tree Giveaways") 

## Fixing the Blue Water Giveaway Data  
neighborhood_intersect <-
  st_intersects(blue_water_giveaway_data, neighborhood_read, sparse = FALSE) |>
  apply(1, any)

county_intersect <-
  st_intersects(blue_water_giveaway_data, baltimore_city_counties, sparse = FALSE) |>
  apply(1, any)

blue_water_giveaway_data <-
  blue_water_giveaway_data |>
   mutate(county = case_when(
    neighborhood_intersect ~ "Baltimore City",
    county_intersect ~ "Baltimore County",
    TRUE ~ "Neither"
  ))

```

```{r Loading Tapestry, eval=FALSE, message=FALSE, include=FALSE}
## Making Tapestry
load('output_data/tap_lu_2024-06-25.RData')

tap_lu |> map(~levels(.)) 

load('output_data/bwb_total_giveaway_with_Tapestry_2019_2024-06-25.RData')

blue_water_giveaway_data_with_Tapestry_2019 |> 
  st_drop_geometry() |> 
  group_by(segment) |> 
  count() |> 
  ggplot(aes(n, segment)) +
  geom_col()
```

```{r Color Palette, eval=FALSE, include=FALSE}
# List of Color Palette
##  (Earth Tones):
   "#f3f0e7"
   "#dbc5a5"
   "#cc9a86"
   "#bd806d"
   "#8c756a"

## Trees (Shades of Green):
   "#2f3456"
   "#5d6e75"
   "#8ba794"
   "#bac6b8"
   "#e9e4dc" 
   
## Branding (Shades of Blue)
   "#008CCC"
   "#005273"
   "#519342"
```
## Location & Time

In the past eleven years of running Blue Water Baltimore's giveaway
program, there has been a total of `r nrow(blue_water_giveaway_data)`
tree giveaways. Although the number is an overestimate since 2017 and
2018 included TreeBaltimore's values, the impact is undeniable. Blue Water's tree plantings have reached homeowners across the state of Maryland, from Frederick to Caroline County.

Figure 1. Giveaway Map of All Jurisdictions
```{r echo=FALSE}
giveaway_layer + state_layer #TODO: Fix the map zoom to focus on Maryland
```

Of course, most of these tree plantings have went to homeowners within the city of Baltimore and Baltimore County. #TODO: Change the verbiage here

Figure 2. Giveaway Heatmap of All Jurisdiction
```{r}
leaflet(blue_water_giveaway_data) |>
  addProviderTiles("CartoDB.Positron") |>
  addHeatmap(lng = ~st_coordinates(blue_water_giveaway_data)[,1], 
             lat = ~st_coordinates(blue_water_giveaway_data)[,2], 
             blur = 5, 
             max = 0.10, 
             radius = 6)
```

More specifically, these giveaways went to northern Baltimore County - near Towson - and northern Baltimore. If you solely look at Baltimore City and Baltimore County, you can see the following trend:

Figure 3. Map of Tree Giveaways per Census Block Group
```{r include=FALSE}
figure_3_census_count_layer <-
  cbg |>
  left_join(
  read_csv("output_data/census_count_year.csv")
    , by = "NAME"
  ) |>
  mutate(count = case_when(
    is.na(count) ~ 0,
    TRUE ~ count
  )) 

figure_3_census_count_legend<- 
  classIntervals(
  figure_3_census_count_layer$`count`,
  n = 5, 
  style = "fisher"
)$brks


figure_3_census_count_layer |>
mapview(zcol = "count", layer.name = "Giveaway Count per Census Block", at = figure_3_census_count_legend)
```

Table 1. Table of Tree Plantings per Jurisdiction
```{r}
table_1_tree_plantings_per_jurisdiction <-
blue_water_giveaway_data |>
  st_drop_geometry() |>
  tabyl(county) |>
  as_tibble() |>
  filter(county %in% c("Baltimore City", "Baltimore County", "Neither"))  |>
  rename(County = county, `Number of Giveaways` = n, Percent = percent) |>
  mutate(Percent = sprintf("%.2f%%", Percent * 100)) 

table_1_tree_plantings_per_jurisdiction |>
DT::datatable()
```

Graph 1. Graph of Tree Plantings per Jurisdiction per Year (Divided by Season)
```{r}
graph_1_season_colors <- c(
  "Fall" = "#008CCC",
  "Spring" = "#005273"
)

graph_1_tree_plantings_year <-
  blue_water_giveaway_data |>
  st_drop_geometry() |>
  group_by(season, year) |>
  summarise(`Number of Giveaways` = n(), .groups = 'drop') |>
  rename(Season = season, Year = year) |>
  mutate(Year = as.double(Year)) |>
  ggplot(aes(x = Year, y = `Number of Giveaways`, fill = Season)) +
   geom_bar(position = "stack", stat = "identity") +
    scale_fill_manual(values = graph_1_season_colors) +
    scale_y_continuous(
      limits = c(0, 2000),         
      breaks = seq(0, 2000, by = 250),  
      labels = scales::comma  
    ) +
    scale_x_continuous (
      breaks = seq(2013, 2023) 
    ) +
    labs(title = "Total Giveaways by Year and Season",
         x = "Year",
         y = "Total Giveaways",
         fill = "Season") +
    theme_minimal() 

ggsave(
  filename = file.path("/Users/eduardosmarin/Desktop/baltimore_tree_partners/deliverables", "location_graph_1_tree_plantings_general.svg"), #TODO: Fix to make more general file saving path
  plot = graph_1_tree_plantings_year, 
  width = 7, 
  height = 7
)

graph_1_tree_plantings_year
```

The amount of tree giveaways that went to either Baltimore City or Baltimore County are nearly split halfway. `r table_1_tree_plantings_per_jurisdiction |> filter(county == "Baltimore City") |> select(Percent) |> pull()` of the tree giveaways went to Baltimore City and `r table_1_tree_plantings_per_jurisdiction |> filter(county == "Baltimore County") |> select(Percent) |> pull()` went to Baltimore County. However, the spread between city and county trees have shifted over the years.

Table 2. Table of Tree Plantings per Jurisdiction per Year
```{r}
table_2_blue_water_baltimore_join <- 
  blue_water_giveaway_data |>
  st_drop_geometry() |>
  tabyl(county, year) |>
  as_tibble() |>
  filter(county %in% c("Baltimore City", "Baltimore County", "Neither"))

 table_2_blue_water_baltimore_join |>
 summarize(across(starts_with("20"), sum, na.rm = TRUE)) |>
 mutate(county = "Total") |>
 bind_rows(table_2_blue_water_baltimore_join) |>
 dplyr::select(11, 1:10) |>
 arrange(match(county, c(setdiff(unique(county), "Total"), "Total"))) |>
 as_tibble() |>
 rename(County = county) |>
 DT::datatable()
```

Graph 2. Graph of Tree Giveaways per Jurisdiction per Year (Divided by Season)
```{r}
graph_2_tree_plantings_year <-
  blue_water_giveaway_data |>
  st_drop_geometry() |>
  group_by(season, year, county) |>
  summarise(`Number of Giveaways` = n(), .groups = 'drop') |>
  rename(Season = season, Year = year, County = county) |>
  mutate(Year = as.double(Year)) |>
  ggplot(aes(x = Year, y = `Number of Giveaways`, fill = Season)) +
   geom_bar(position = "stack", stat = "identity") +
    scale_fill_manual(values = graph_1_season_colors) +
    scale_y_continuous(
      limits = c(0, 2000),         
      breaks = seq(0, 2000, by = 250),  
      labels = scales::comma  
    ) +
    scale_x_continuous (
      breaks = seq(2013, 2023) 
    ) +
    labs(title = "Total Giveaways by Year and Season",
         x = "Year",
         y = "Total Giveaways",
         fill = "Season") +
    theme_minimal() +
    facet_wrap(~ County)

ggsave(
  filename = file.path("/Users/eduardosmarin/Desktop/baltimore_tree_partners/deliverables", "location_graph_2_tree_plantings_jurisdiction.png"), #TODO: Fix to make more general file saving path
  plot = graph_2_tree_plantings_year, 
  width = 21, 
  height = 7
)

graph_2_tree_plantings_year
```

The table below will highlight the amount per year and season but a larger scale to understand the magnitude of the previous graph.

Table 3. Table of Plantings in Each Neighborhood
```{r echo=FALSE}
  blue_water_giveaway_data |>
  st_drop_geometry() |>
  tabyl(year, season) |>
  as_tibble() |>
  add_row(year = "2020", Fall = 0, Spring = 0) |>
  mutate(Total = Fall + Spring) |>
  arrange(desc(Total)) |>
  rename(Year = year)
  DT::datatable()
```

The following section will take a more in depth look at the distribution of tree plantings in Baltimore, specifically at the neighborhood level. These will show the final destinations of those within the city:

Figure 4. Map of Tree Plantings in Baltimore Neighborhoods

```{r echo=FALSE}
figure_4_tree_neighborhoods_layer <-
neighborhood_read |>
  left_join(
    read_csv("output_data/neighborhood_count_bwb.csv")
    , by = "Name"
  ) |>
  mutate(count = case_when(
    is.na(count) ~ 0,
    TRUE ~ count
  )) |>
  rename(`Number of Trees` = count) |>
  mutate(Density = `Number of Trees` / `Area (Square Miles)`) 

figure_4_tree_neighborhood_legend <- 
  classIntervals(
  figure_4_tree_neighborhoods_layer$`Density`,
  n = 10, 
  style = "fisher"
)$brks

figure_4_tree_neighborhoods_layer |>
  mapview(zcol = "Density", layer.name = "Giveaway Density (Tree per Square Mile)", at = figure_4_tree_neighborhood_legend) 
```

Table 4. Table of Plantings in Each Neighborhood
```{r echo=FALSE, message=FALSE}
 figure_4_tree_neighborhoods_layer |>
  st_drop_geometry() |>
  dplyr::select("Name","Density") |>
  arrange(desc(`Density`)) |>
  DT::datatable()
```

These may come surprising at first. However, these numbers are taking into consideration the amount of space within each jurisdiction. They are your number of trees in each neighborhood divided by the total square miles of each neighborhood. The following highlights the absolute value of tree plantings in each neighborhood. 

Table 5. Table of Plantings in Each Neighborhood
```{r echo=FALSE, message=FALSE}
figure_4_tree_neighborhoods_layer |>
  st_drop_geometry() |>
  dplyr::select("Name","Number of Trees") |>
  arrange(desc(`Number of Trees`)) |>
  DT::datatable()
```

Regardless, another interesting aspect to consider is how have the absolute number of tree plantings have changed over the years. The decision to not observe the density of tree plantings is since neighborhood boundaries can change, alongside the amount of plantable space over time.

Graph 3. Graph of Tree Plantings in Each Neighborhood Over the Years (Top 20)
```{r}
graph_3_filter_neighborhoods <-
  figure_4_tree_neighborhoods_layer |>
  dplyr::select(Name, `Number of Trees`) |>
  arrange(desc(`Number of Trees`)) |>
  slice(1:10)

graph_3_tree_plantings_year_neighborhood <-
figure_4_tree_neighborhoods_layer |>
  st_drop_geometry() |>
  left_join(
    read_csv("output_data/neighborhood_count_year_bwb.csv"), 
    by = "Name"
  ) |>
  rename(Year = year, Neighborhood = Name) |>
  dplyr::select(Neighborhood, `Number of Trees` = count, Year) |>
  dplyr::filter(Neighborhood %in% graph_3_filter_neighborhoods$Name) |>
  pivot_wider(names_from = Neighborhood, values_from = `Number of Trees`) |>
  add_row(Year = 2020) |>
  arrange((Year)) |>
  pivot_longer(cols = c("Ednor Gardens-Lakeside", "Hamilton", "Hampden", "Hamilton Hills", "Homeland", "Lauraville","Mount Washington","Roland Park","Westfield","Charles Village"), names_to = "Neighborhood", values_to = "Number of Trees") |>
  mutate(Neighborhood = factor(Neighborhood, levels = graph_3_filter_neighborhoods$Name)) |>
  ggplot(aes(x = Year, y = `Number of Trees`)) +
  geom_bar(position = "dodge", stat = "identity", fill = "#519342") +
  scale_x_continuous(breaks = seq(2013, 2023)) +
  coord_flip() +
  labs(title = "Top 10 Neighborhoods and Number of Trees", x = "Year", y = "Number of Trees") +
  theme_minimal() +
  facet_wrap(~ Neighborhood) 

ggsave(
  filename = file.path("/Users/eduardosmarin/Desktop/baltimore_tree_partners/deliverables", "location_graph_3_tree_plantings_neighborhoods_year.png"), #TODO: Fix to make more general file saving path
  plot = graph_3_tree_plantings_year_neighborhood, 
  width = 21, 
  height = 10
)
```

As you can see, the top ten neighborhoods, in descending order, are `r paste(graph_3_filter_neighborhoods$Name, collapse = ", ")`. #TODO: Add oxford "," The most surprising feature is that these neighborhoods vary greatly in terms of when their largest contribution were. Some, like Ednor Gardens-Lakeside, had the most giveaways in 2013, while others had them more recently like Hamilton Hills. 

Nevertheless, as important as it is to understand when and where these plantings have gone, it is also important to understand where they have been received.

## Provider
```{r include=FALSE}
unique_provider <-
 blue_water_giveaway_data |>
  st_drop_geometry() |>
  tabyl(gvwy_lc) |> 
  as_tibble() |> 
  arrange(desc(n)) |>
  slice(-1) 

unique_provider_na <-
 blue_water_giveaway_data |>
  st_drop_geometry() |>
  tabyl(gvwy_lc) |> 
  as_tibble() |> 
  arrange(desc(n)) |>
  slice(1) 
```

In terms of giveaway location, there are approximately `r nrow(unique_provider)` different giveaway locations over the last ten years. Each of them vary, especially in terms of amount.

Table 6. Table of Giveaway Locations 
```{r echo=FALSE}
  blue_water_giveaway_data |>
  st_drop_geometry() |>
  tabyl(gvwy_lc) |> 
  rename(`Giveaway Location` = gvwy_lc) |>
  as_tibble() |> 
  arrange(desc(n)) |>
  slice(-1) |> 
  rename("Number of Trees" = n, Percent = percent) |>
  mutate(Percent = round(100 * Percent, 2)) |>
  dplyr::select(-`valid_percent`) |>
  DT::datatable()
```

In the case of the amount of tree giveaway locations, the most stems from Stillmeadow Community Fellowship. However, please understand that the largest majority are undefined, totaling `r unique(unique_provider_na$n)`.

Graph 4. Graph of Total Giveaway Locations 
```{r echo=FALSE}
graph_4_giveaway_locations <-
  blue_water_giveaway_data |>
  st_drop_geometry() |>
  tabyl(gvwy_lc) |>
  as_tibble() |> 
  arrange(desc(n)) |>
  slice(-1) |> 
  rename("Number of Trees" = n, `Giveaway Location` = gvwy_lc) |>
  mutate(percent = round(100 * percent, 2)) |>
  ggplot(aes(x = reorder(`Giveaway Location`, `Number of Trees`), y = `Number of Trees`)) +
    geom_col(fill = "#519342") +
    coord_flip() +
    labs(title = "Giveaway Location and Number of Trees", x = "Giveaway Location", y = "Number of Trees")
    theme_minimal() +
    NULL

# TODO: Fix the graph (currently offset)    
# ggsave(
#   filename = file.path("/Users/eduardosmarin/Desktop/baltimore_tree_partners/deliverables", "provider_graph_4_tree_plantings_neighborhoods_year.png"), #TODO: Fix to make more general file saving path
#   plot = graph_4_giveaway_locations, 
#   width = 21, 
#   height = 10
# )
```

The following... #TODO: Find more graphs to do centering on giveaway locations. Potentially removing this section and just having it for the origin-destination mapping.

## Species
```{r include=FALSE}
colors_species <- c(
  "Serviceberry" = "#9e0142",
  "Sweetbay Magnolia" = "#d53e4f", 
  "Pawpaw" = "#f46d43",
  "American Persimmon" = "#fdae61", 
  "Common Fig" = "#fee08b",
  "American Plum" = "#e6f598", 
  "American Holly" = "#abdda4",
  "Black Gum" = "#92b1e2",
  "Eastern Redbud" = "#3288bd", 
  "American Hornbeam" = "#5e4fa2"
)
```

Species information is as important as location, timing, and provider. The survey highlight 

Figure 4. Map of Top 10 Species Giveaway
```{r echo=FALSE}
blue_water_giveaway_data |>
  mutate(cmmn_nm = case_when(
    cmmn_nm %in% c("Serviceberry","Sweetbay Magnolia","Pawpaw","American Persimmon","Common Fig","American Plum","American Holly","Black Gum","Eastern Redbud", "American Hornbeam") ~    cmmn_nm, 
    TRUE ~ NA_character_
  )) |>
  filter(!is.na(cmmn_nm)) |>
  dplyr::select(year, season, cmmn_nm, gns_spc) |>
  mapview(zcol = "cmmn_nm", col.regions = colors_species, layer.name = "Top 10 Tree Species Planted from Blue Water Baltimore")

```

#### Common Name Datatable

```{r echo=FALSE}
## making the species dataframe
  blue_water_giveaway_data  |>
  st_drop_geometry() |>
  tabyl(common_name) |>
  arrange(desc(n)) |>
  mutate(percent = round(100 * percent, 2)) |>
  DT::datatable()
```

#### **Common Name Graph (Both Rare and Common)**

```{r echo=FALSE}
## creating the groups for the graph
  blue_water_giveaway_data  |>
  st_drop_geometry() |>
  tabyl(common_name) |>
  arrange(desc(n)) |>
  mutate(percent = round(100 * percent, 2)) |>
  mutate(group = rep(c("common", "rare"), each = 42)) |>
  ggplot(aes(n, reorder(common_name, n))) + # pushes the bars in the desc order
  geom_col() +
  theme_bw(10) +
  labs(y = 'Common Name'
       , x = 'Number of trees'
       , title = 'Which tree species were most and least common?'
       , subtitle = paste0('lots of Serviceberry... ')
       ) +
  facet_wrap(~group, scales = 'free') +
  NULL

```


```
